---
title: "Mixed Effects Logistic Regression"
author: "Alex LoPilato"
output: 
  ioslides_presentation:
    widescreen: true
    mathjax: default
    css: "style.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../R/01-dependencies.R", local = knitr::knit_global())
source("../R/02-functions.R", local = knitr::knit_global())
source("../R/03-simulated-data.R", local = knitr::knit_global())
source("../R/04-model-fitting.R", local = knitr::knit_global())
```

## <font size = "6">Overview</font>

- Recap of LMERs
- Primer on Logistic Regression
- Introduction to GLMERs 
- Model Fitting Example 

## <font size = "6">Linear Mixed Effects Regression Model Primer</font>

- A flexible class of linear models that allow one to adjust for and model dependencies that occur in one's data because of how the data were generated. 

- Linear Mixed Effects Regression Models (LMERs) allow us to relax the assumption that our data are independent of one another. 

## <font size = "6">Dependency by Design</font> {.smaller}
Two common designs that can cause dependency:

1. Nested Design 
\tiny
    ```{r, echo = FALSE}
    Employee <- 1:4
    Firm <- rep(1:2, each = 2)
    xtabs(~ Employee + Firm)
    ```
\normalsize
2. Crossed Design
\tiny
    ```{r}
    Employee <- rep(1:3, 2)
    Rater <- rep(1:2, each = 3)
    xtabs(~Employee + Rater)
    ```

## Random Effects
$$Var(Y_{ij}) = \sigma^2 + \tau^2 $$

## <font size = "6">Logistic Regression Model Primer</font> 

## <font size = "6">Interpreting Logistic Regression Coefficients</font>

## <font size = "6">Plotting Predicted Probabilities</font>

## <font size = "6">Mixed Effects Logistic Regression</font>

## <font size = "6">Mixed Effects Logistic Regression</font>

## <font size = "6">Motivating Example</font>

## <font size = "6">Cross Classified Logistic Regression Model</font> 

## <font size = "6">Model Fitting Strategy</font>

- The Top-Down Strategy 
- The Step-Up Strategy 

## <font size = "6">Empty Model</font>

## <font size = "6">Adding Main Effects of Predictors</font>

## <font size = "6">Adding Interaction Effects</font>

## <font size = "6">Assessing Overall Model Fit</font>

## <font size = "6">Interpreting the Final Model</font>


